---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r}
unzip("activity.zip")
data <- read.csv("activity.csv")
```

## What is mean total number of steps taken per day?

1. Make a histogram of the total number of steps taken each day

```{r, warning=FALSE}
total_steps <- aggregate(list(Total = data$steps), by=list(Date = data$date), FUN=sum, na.rm=TRUE)
hist(total_steps$Total)
``` 

2. Calculate and report the mean and median total number of steps taken per day

```{r, warning=FALSE}
mean_steps <- aggregate(list(Mean = data$steps), by=list(Date = data$date), FUN=mean, na.rm=TRUE)
mean_steps
``` 

```{r, warning=FALSE}
median_steps <- aggregate(list(Median = data$steps), by=list(Date = data$date), FUN=median, na.rm=TRUE)
median_steps
```

## What is the average daily activity pattern?

1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r}
x_axis_values <- data[data$date=="2012-10-04",3]
mean_steps_by_interval <- aggregate(list(mean = data$steps), by=list(interval = data$interval), FUN=mean, na.rm=TRUE)
plot(x_axis_values, mean_steps_by_interval$mean, type="l")
```

2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
x_axis_values[which.max(mean_steps_by_interval$mean)]
```

## Imputing missing values

1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

```{r}
sum(is.na(data$steps))
```

2. Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.

*For missing step values, a value will be set that is equivalent to the mean value for that interval as taken across all days for that interval*

3. Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r}
interpreted_data <- data

for (i in 1:length(interpreted_data$steps)) {
  
  if(is.na(interpreted_data[i,]$steps)){
      interpreted_data[i,]$steps = mean_steps_by_interval[mean_steps_by_interval$interval == interpreted_data[i,]$interval,2]
    }
}

```

4. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day.
```{r}
total_interpreted_steps <- aggregate(list(Total = interpreted_data$steps), by=list(Date = data$date), FUN=sum, na.rm=TRUE)
hist(total_interpreted_steps$Total)
```

```{r, warning=FALSE}
interpreted_mean_steps <- aggregate(list(Mean = interpreted_data$steps), by=list(Date = interpreted_data$date), FUN=mean, na.rm=TRUE)
interpreted_mean_steps
``` 

```{r, warning=FALSE}
interpreted_median_steps <- aggregate(list(Median = interpreted_data$steps), by=list(Date = interpreted_data$date), FUN=median, na.rm=TRUE)
interpreted_median_steps
```

## Are there differences in activity patterns between weekdays and weekends?

1. Create a new factor variable in the dataset with two levels -- "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.

```{r}

if(!require(timeDate)){
  install.packages("timeDate")
}
library(timeDate)

interpreted_data$isWeekday = as.factor(ifelse(isWeekday(interpreted_data$date), "weekday", "weekend"))
```

2. Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).

```{r}
mean_steps <- aggregate(list(mean = interpreted_data$steps), by=list(interval = interpreted_data$interval, isWeekday = interpreted_data$isWeekday), FUN=mean, na.rm=TRUE)

if(!require(lattice)){
  install.packages("lattice")
}
library(lattice)

xyplot(mean ~ interval | factor(isWeekday), data = mean_steps, type="l", layout=c(1,2))

```
